Why do we Eval?\\[2mm]\Large A large-scale study of Eval usage in R

### 2020-11-26 inst/paper/main.tex
intro

application domains.  It is thus reasonable to wonder if the results from
previous work can be carried over to other contexts. For this paper, we have
picked the data science language R~\cite{r} as our object of study. R is
both different from JavaScript in linguistic terms and in terms of its user
base. JavaScript was designed to run untrusted code in browser, while R was
designed for statistical computing on a desktop. JavaScript is used as a
general purpose language by a wide community of programmers; while R is used
for scientific computing by data scientists and domain experts with, often,
limited programming experience.

This paper sets out to highlight the differences in usage patterns of \eval
between these two languages and user communities. Our results are broadly
applicable as they increase our understanding of the range of uses of \eval,
present new use-cases, and, hopefully, provide insights into reasonable
under-approximations. More narrowly, for researchers interested in static
analysis of data science code, we provide a publicly available corpus of
programs and a toolchain that can be customized to deliver fine-grained
information about the usage and impact of \eval.

@@

Our corpus has been constructed to reflect the different level of
sophistication in the R community. We distinguish between the core
implementers of the language, developers with both extensive programming
experience and keen understanding of the language semantics, library
implementers, developers with some programming experience and a working
knowledge of R, and end-users, who are typically not expert programmers and
often have only a cursory knowledge of the language.\footnote{To illustrate
  the above consider that R uses lazy evaluation like Haskell. The authors
  have informally surveyed end-users, including computer scientists, and did
  not find a single one aware of this fact. Library developers, on the other
  hand, know about laziness and program defensively around it.}  Thus, our
corpus distinguished between \emph{core packages} (there are
\CorpusCorePackages such packages that are distributed together with the R
virtual machine), \emph{CRAN packages} (we selected \CorpusPackages curated
packages that pass stringent community quality checks and are equipped with
tests and sample data), \emph{Kaggle scripts} (we used \CorpusFinishedKaggle
end-user written programs that performs a particular data analysis
task). There are reasons to believe that \eval usage patterns may differ
between these datasets. The core R code is extremely stable, it has only
been maintained for the last 20 years with very few far reaching
changes. The libraries represent a more lively ecosystem with new libraries
added each day.  Finally, end user code is often one-use scripts thrown
together and sometimes never revisited.

>>>>

\vspace{2mm}\noindent\emph{The here and now.}  In this study, we
investigate the usage of \eval in programs written in the R
programming language. R is language designed by statisticians for
applications in data science~\cite{r}. What makes looking at R after
JavaScript interesting is that, while both languages are dynamic, they
are quite different. While one can program in an object-oriented style
in R, like in JavaScript, R is primarily a lazy, untyped, functional
language.  JavaScript was designed to run untrusted code in browser,
while R is used for statistical computing on desktops. JavaScript is a
general purpose language used by a wide community of programmers;
while R is used for scientific computing by data scientists and domain
experts with, often, limited programming experience. One can
distinguish between library implementers, developers with some
programming experience and a working knowledge of R, and end-users,
who are typically not expert programmers and often have only a cursory
knowledge of the language.\footnote{Consider that R evaluates function
argument lazily, just like Haskell. We informally surveyed end-users,
including computer scientists, and did not find a single user aware of
this fact. Library developers, on the other hand, know about laziness
and program defensively around it.} Our goal is thus to highlight the
differences in usage between JavaScript and R, and try to explain
those differences in terms of language features, application domain
and programmer experience.

\vspace{2mm}\noindent\emph{Our contributions.} 
TODO

@@

Our corpus has been constructed to reflect the different level of
sophistication in the R community.

Thus, our
corpus distinguished between \emph{core packages} (there are
\CorpusCorePackages such packages that are distributed together with the R
virtual machine), \emph{CRAN packages} (we selected \CorpusPackages curated
packages that pass stringent community quality checks and are equipped with
tests and sample data), \emph{Kaggle scripts} (we used \CorpusFinishedKaggle
end-user written programs that performs a particular data analysis
task). There are reasons to believe that \eval usage patterns may differ
between these datasets. The core R code is extremely stable, it has only
been maintained for the last 20 years with very few far reaching
changes. The libraries represent a more lively ecosystem with new libraries
added each day.  Finally, end user code is often one-use scripts thrown
together and sometimes never revisited.
